---
title: "Hands-on Exercise 1: Geospatial Data Wrangling with R"
execute: 
  warning: false
date: "`r Sys.Date()`"
---

# Getting Started

## Installing and loading R packages

In this section, I will install and load tidyverse and sf packages.

```{r}
pacman::p_load(tidyverse, sf)

```

# Importing a geospatial data in shapefile format

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")

```

## Importing polyline feature data in shapefile form

```{r}
#| echo: false
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPathGazette")

```

## Importing GIS data in kml format

```{r}
#| echo: false
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")

```

# Checking the Content of A Simple Feature Data Frame

##Working with st_geometry()

```{r}
#| echo: false
st_geometry(mpsz)

```

##Working with glimpse()

```{r}
#| echo: false
glimpse(mpsz)

```

##Working with head()

```{r}
#| echo: false
head(mpsz, n=5) 

```

# Plotting the Geospatial Data

```{r}
#| echo: false
plot(mpsz)
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```

#Working with Projection

##Assigning EPSG code to a simple feature data frame

```{r}
#| echo: false
st_crs(mpsz)

```

```{r}
#| echo: false
mpsz3414 <- st_set_crs(mpsz, 3414)
```

```{r}
#| echo: false
st_crs(mpsz)

```

##Transforming the projection of preschool from wgs84 to svy21.
```{r}
#| echo: false
preschool3414 <- st_transform(preschool, 
                              crs = 3414)

```

#Importing and Converting An Aspatial Data

## Importing the aspatial data
```{r}
#| echo: false
listings <- read_csv("data/aspatial/listings.csv")

```

```{r}
#| echo: false
list(listings) 
```

##Creating a simple feature data frame from an aspatial data frame
```{r}
#| echo: false
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
#| echo: false
glimpse(listings_sf) 
```

#Geoprocessing with sf package

##Buffering

```{r}
#| echo: false
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

```{r}
#| echo: false
buffer_cycling$AREA <- st_area(buffer_cycling)
```


```{r}
#| echo: false
sum(buffer_cycling$AREA)
```

##Point-in-polygon count

```{r}
#| echo: false
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

```{r}
#| echo: false
summary(mpsz3414$`PreSch Count`)
```

```{r}
#| echo: false
top_n(mpsz3414, 1, `PreSch Count`)
```


The solution:

Firstly, the code chunk below uses st_area() of sf package to derive the area of each planning subzone.
```{r}
#| echo: false
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```
Next, mutate() of dplyr package is used to compute the density by using the code chunk below.
```{r}
#| echo: false
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

#Exploratory Data Analysis (EDA)

```{r}
#| echo: false
hist(mpsz3414$`PreSch Density`)
```

```{r}
#| echo: false
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```
DIY: Using ggplot2 method, plot a scatterplot showing the relationship between Pre-school Density and Pre-school Count.

The solution:
```{r}
#| echo: false
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```